Notes on generating initial models for mapping into {\tt MAESTRO}

### 2008-08-27 docs/initial_models/paper.tex
now reflects what's in cvs.  I'm much happier with this algorithm, it's very clean compared to what was there before.  All the test problems seem to perform well.

For the finest level, we can set the initial data equal to the model data.  For the non-finest levels, we use linear interpolation to compute the initial data.  For cells that are covered by finer cells, we average $\rho^{\init}, (\rho h)^{\init},$ and $(\rho X)^{\init}$ from the fine cell to get the coarse cell value.  Then, for every cell not at the finest level, we call the EOS to compute $p^{\init},T^{\init} = p^{\init},T^{\init}(\rho^{\init},h^{\init},X^{\init})$.\MarginPar{Now $p^{\init}$ and $\rho^{\init}$ are not in HSE.  Is this a problem?}
Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.  We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.  Here are the steps:
\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.  Note that the multifab version of $p$ is temporary, and can be deallocated after this step.
\item Perturb $T$, then compute $\rho = \rho(p^{\init},T,X)$.
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
Refer to figure \ref{Fig:Multilevel Initialization}.  The dots represent the initial data, which was generated at the finest resolution on the domain, and is in HSE as well as thermodynamic equilibrium.  The circles represent the fine level data points in the 1D array.  The squares represent the coarse level data points in the 1D array.  We modify step 1 in \S~\ref{Sec:Creating the Base State and Full State} as follows:
\item Step 1
\item Copy the initial data from the dots directly into the circles.  For the squares, compute only $\rho$ and $X$ by using linear interpolation from the two nearest corresponding dots.
\item Integrate to get $p$ in the squares.  We need the fancy stencils at coarse-fine interfaces in equations (\ref{Coarse-Fine Stencil}) and (\ref{Fine-Coarse Stencil}).  Away from coarse-fine interfaces, we can use our standard HSE discretization.
\item Call EOS to get $T$ and $h$ in the squares, i.e., $T,h = T,h(\rho,p,X)$.
\item We now copy this 1D profile onto the full multifab.

@@

\caption{Multilevel Initialization}

@@

\subsubsection{Coarse-Fine HSE Discretization}

>>>>

We do the same thing as the single-level case except that we generate initial data at the non-finest levels by using linear interpolation from the two nearest model points (see Figure \ref{Fig:Multilevel Initialization}).  Note that the initial data is not in HSE nor is it thermodynamically consistent on non-finest cells.  We will enforce these on the base state and full state later.

@@

\caption{Multilevel Initialization.  The data from the initial model is represented by the dots on the right.  The initial data at the finest level is represented by the circles.  The initial data at non-finest levels is represented by the squares.  We copy data from the dots directly into the circles.  We use linear interpolation using the two nearest points to copy data from the dots to the squares.}

@@

Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.  We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is thermodynamically consistent with $p_0$.  Here are the steps:
\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.  Note that this particular multifab version of $p$ is temporary, and can be deallocated after this step.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Set $(\rho h)_0 = \overline{(\rho h)}$.
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
\item If {\tt perturb\_model} = F, skip the next two steps and instead set $\rho_0 = \rho^{\init}$.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.  Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.  Use the coarse-fine interface discretizations given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
\item Set $(\rho h)_0 = \overline{(\rho h)}$.
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
\subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
